Storm surges in the Strait of Georgia simulated with a regional model

### 2015-05-14 AO_guides/manuscriptAO.tex
Finished constitency and checked for typos

Sea surface height anomaly (or remote forcing) was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain (Figure \ref{fig:domain}).  A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.  For the storm surge simulations, hourly variations in sea surface height were used.  These values are the Tofino tide gauge values with the tides removed. {\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge for 67 constituents. In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes. Long period constituents such as SA, SSA, etc are not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends. Further, constituents with a low signal to noise ratio were neglected. } The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.

@@

Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range. {\color{red} As such, model water level predictions were corrected by accounting for this error. Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations. As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.  The difference between these two predictions is added to the model ses surface height as a correction. Note that the harmonic analysis to generate the constituents is applied in the same manner as that used in the forcing. That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed. The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals. All times are reported in UTC.

@@

{\color{red} This case was also used to asses the importance of the northern boundary residual forcing and was used to brielfy examine stratified effects. When the northern boundary forcing is exlcuded the surge amplitude at Campbell River drops by about 10 cm. The other stations were not very much affected. This was also observed in another hindcast (Dec 2006). The effect of stratification was examined by considering a case with an initially horizontally uniform stratification and comparing to the fully stratified case. The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period. Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic mdoel as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}. We plan to investigate stratified effects in more detail in a future study.}

@@

First, it is clear that remote forcing at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}). When the remote forcing is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.  A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.  When local atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in the remote forcing.

%Local winds
The contribution to the surge due to local atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds. Removal of the local atmospheric forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River. By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay. The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model. A simulation with only tides and local wind forcing produces a surge of 7.8~cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.  Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.

@@

Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}). The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the difference between the modelled residual and surge-only simulation (Figure \ref{fig:tidesurge_spatial}) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation. However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, performing a surge-only simulation and adding the predicted tides \textit{a posteriori} could be more computationally efficient with only a small decrease in the accuracy of the prediction.

@@

\subsection{Spatial Extent of Surge}\label{sec:spatial}
A major wind storm on Dec 15, 2006 is hindcasted next. Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}. Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude. First, an assessment of the accuracy of the hindcast is provided.

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}). The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals. Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain). {\color{red} Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}

@@

The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC. The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain. Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70} cm, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60} cm {\color{red} (Figure \ref{fig:spatial})}. Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast. At about 11:30 UTC, the winds at Point Atkinson switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.

@@

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.  {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia. Note that since local winds are stronger in this case, the local wind contribution is more significant than in the Feb 2006 case (6 cm contribution at Point Atkinson). However, this is still les than 10 \% of the total surge amplitude. }

@@

\subsection{Strong Wind Event}\label{sec:wind}
Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered. {\color{red} Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.}  Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}). Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20. {\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded. Although {\color{red} the sea surface height plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.

\citet{danard2003storm} discuss the role of wind fetch, the distance over which wind acts on the surface of a water basin, in setting up seiche motion in lakes and its contribution to storm surges. As wind blows over the surface of a lake, the water level becomes elevated on the downwind side and depressed on the upwind side in response to the wind-induced currents. The longer the fetch, the more significant the tilting of the water surface. These changes in water level can also contribute to storm surges.  During the high wind events on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin. In fact, the surge amplitude at Victoria on Nov 20 was 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm). Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson. The model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.

@@

The modelled water levels and residuals are too high for this hindcast, except at Victoria (Table \ref{tab:statistics}). Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations. Generally, the skills scores are very good, scoring above 0.91 at all stations.

@@

First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson and at Sandheads, a lighthouse station in the southern Strait of Georgia. For each of the models, the wind speed and direction at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison. Although the low resolution CGRF model has a grid spacing of 33 km, two grid points align closely with these locations. At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012. At Sandheads, both the low resolution and high resolution models have good agreement with observations. The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water. Although not shown here, both models have a similar sea level atmospheric pressure at these two locations. Note that data assimilation is used in the parent model of the nested HRDPS product.

@@

Hindcasts with both atmospheric forcing products agree well with the observations (Figure \ref{fig:dec2012_surge}). Willmott skill scores for both modelled residuals are above 0.92 at these four locations. Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the {\color{red}strong} wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher. Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm. Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm. Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products. Much of the fine scale detail offered by the high resolution atmospheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind. This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain; however, local wind patterns contribute to wave conditions which can impact flooding and are not accounted for in this model.

@@

Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated. In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally higher skill scores for the water level heights.  Further, in a single simulation, the skill scores are fairly consistent between the four stations examined. We have also assessed the relative importance of different forcing factors contributing to the storm surge. The sea surface height anomaly entering the system from the Pacific Ocean, or remote forcing,  contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}. Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude. However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season. {\color{red}Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.} Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands. The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.

@@

Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}. Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics. Additionally, a biological-chemical modelling component will be added.

>>>>

Sea surface height anomaly (or remote forcing) was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain (Figure \ref{fig:domain}).  A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.  For the storm surge simulations, hourly variations in sea surface height were used.  These values are the Tofino tide gauge values with the tides removed. {\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge. A harmonic analysis was performed for 67 constituents. In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes. Long period constituents such as SA, SSA, etc were not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends. Further, constituents with a low signal to noise ratio were neglected. } The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.

@@

Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range. {\color{red} As such, model water level predictions were corrected by accounting for this error. Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations. As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.  The difference between these two predictions is added to the model sea surface height as a correction. Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing. That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed. The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals. All times are reported in UTC.

@@

{\color{red} This case was also used to asses the importance of the northern boundary residual forcing and was used to brielfy examine stratified effects. When the northern boundary forcing is exlcuded the surge amplitude at Campbell River drops by about 10 cm. The other stations were not very much affected. This was also observed in another hindcast (Dec 2006). The effect of stratification was examined by considering a case with an initially horizontally uniform stratification and comparing to the fully stratified case. The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period. Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic model as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}. We plan to investigate stratified effects in more detail in a future study.}

@@

First, it is clear that remote forcing at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}). When the remote forcing is not included as a forcing condition, the surge drops from a maximum {\color{red}84 cm} to 11 cm at Point Atkinson and its maximum occurs eight hours later.  A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.  When local atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in the remote forcing.

%Local winds
The contribution to the surge due to local atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds. Removal of the local atmospheric forcing leads to a drop of about {\color{red}2 cm} in the maximum surge at Point Atkinson and {\color{red}2.5 cm} at Campbell River. By contrast, the surge amplitude increases by {\color{red} 4.5 cm} at Victoria and 3.6 cm at Patricia Bay. The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model. A simulation with only tides and local wind forcing produces a surge of 7.8~cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.  Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.

@@

Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}). The most significant differences occur at Point Atkinson where the maximum of the surge-only case is {\color{red} 8.4 cm} higher than maximum modelled residual, about 10\% of the total surge amplitude.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the difference between the modelled residual and surge-only simulation (Figure \ref{fig:tidesurge_spatial}) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation. However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, performing a surge-only simulation and adding the predicted tides \textit{a posteriori} could be more computationally efficient with only a small decrease in the accuracy of the prediction.

@@

\subsection{Spatial Extent of Surge}\label{sec:spatial}
A major wind storm on Dec 15, 2006 is hindcasted next. Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached {\color{red}94 cm}, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}. Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude. First, an assessment of the accuracy of the hindcast is provided.

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above {\color{red}0.89} and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}). The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals. Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain). {\color{red} Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}

@@

The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC. The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain. Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70 cm}, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60 cm} {\color{red} (Figure \ref{fig:spatial})}. Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast. At about 11:30 UTC, the winds at Point Atkinson switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.

@@

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.  {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia. Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case. However, this is still less than 10 \% of the total surge amplitude. }

@@

\subsection{Strong Wind Event}\label{sec:wind}
Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered. {\color{red} Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.}  Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}). Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about {\color{red}55 cm} at Point Atkinson, and a second, larger surge of {\color{red} 70 cm} on Nov 20. {\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded. Although {\color{red} the sea surface height plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.

\citet{danard2003storm} discuss the role of wind fetch, the distance over which wind acts on the surface of a water basin, in setting up seiche motion in lakes and its contribution to storm surges. As wind blows over the surface of a lake, the water level becomes elevated on the downwind side and depressed on the upwind side in response to the wind-induced currents. The longer the fetch, the more significant the tilting of the water surface. These changes in water level can also contribute to storm surges.  During the high wind events on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin. In fact, the surge amplitude at Victoria on Nov 20 was {\color{red}71.8 cm} which is slightly higher than that observed at Point Atkinson {\color{red}(70 cm)}. Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson. The model results do not reflect a higher surge at Victoria {\color{red}(54 cm versus 58.7~cm at Point Atkinson)}, perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.  {\color{red} There is a fairly significant delay in the timing of the maximum surge at most stations (Table \ref{tab:peak}). The model's surge curve is much broader and does not contain the narrow peak on Nov 20 as seen in the observations.}

@@

The modelled water levels and residuals are, {\color{red} on average,} too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).  Generally, the skills scores are very good, scoring above {\color{red} 0.89} at all stations.

@@

{\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.  Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and direction {\color{red} the wind is heading} at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison. Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations. At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012. At Sandheads, both the low resolution and high resolution models have good agreement with observations. The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water. Although not shown here, both models have a similar sea level atmospheric pressure at these locations. Note that data assimilation is used in the parent model of the nested HRDPS product.

@@

Hindcasts with both atmospheric forcing products agree well with the observations (Figure \ref{fig:dec2012_surge}). Willmott skill scores for both modelled residuals are above {\color{red} 0.84} at these four locations. {\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}.} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.  The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).} Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm. Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was {\color{red}66.3 cm}. Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products. Much of the fine scale detail offered by the high resolution atmospheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind. This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain; however, local wind patterns contribute to wave conditions which can impact flooding and are not accounted for in this model.

@@

Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated. In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than {\color{red}0.84} in all cases and generally higher skill scores for the water level heights.  Further, in a single simulation, the skill scores are fairly consistent between the four stations examined. We have also assessed the relative importance of different forcing factors contributing to the storm surge. The sea surface height anomaly entering the system from the Pacific Ocean, or remote forcing,  contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}. Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude. However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season. {\color{red}Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.} Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands. The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.

@@

Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}. Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics {\color{red} and investigating stratified effects on surge amplitudes.} Additionally, a biological-chemical modelling component will be added.
